[Re] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model

### 2018-02-06 article/ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.tex
Update version of figures and manuscript

\documentclass[10pt,a4paper,onecolumn]{article}
% * <aroquesilva@gmail.com> 2017-09-19T13:55:25.418Z:
%
% ^.
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\usepackage{cleveref} %added by Renan
\usepackage{multirow} %added by Rodrigo
\usepackage{amsmath,amssymb} %added by Roque
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Renan O. Shimoura,
      Nilton L. Kamiji,
      Rodrigo F.O. Pena,
  },
            pdftitle=
{
[Re] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}

@@

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.bib}

@@

Network neurons are described by the leaky integrate-and-fire (LIF) neuron model. The subthreshold membrane voltage of neuron $i$ obeys the equation
\begin{equation}
\label{Eq:LIF}
\dot{V}_i(t) = -\left( V_i(t) - V_{\rm reset} \right) /\tau_m + I_i(t)/C_m,
\end{equation}
where $\tau_m$ is the membrane time constant, $C_m$ is the membrane capacitance, $V_{\rm reset}$ is the reset potential, and $I_i(t)$ is the total input current. When $V_i(t) \geq V_{\rm th}$ the neuron emits a spike and the voltage is reset to $V_{\rm reset}$, remaining fixed at $V_{\rm reset}$ for a refractory period $\tau_{\rm ref}$. The total input current $I_i(t)$ is divided into external $I_{i, \rm ext}(t)$ and synaptic $I_{i, \rm syn}(t)$. Whenever an excitatory (or inhibitory) neuron $j$ presynaptic to neuron $i$ fires at time $t_{j}^{f}$, the synaptic current to neuron $i$ changes at time $t_{j}^{f}+d_e$ (or $d_i$) by an amount $w$ (or $-g w$), where $w$ is the excitatory synaptic weight, $g$ is the relative weight of the inhibitory synapse, $d_e$ is the excitatory synapse transmission delay, and $d_i$ is the inhibitory synapse transmission delay. In the absence of synaptic inputs the synaptic current changes as
\begin{equation}
\label{Eq:Isyn}
\dot{I}_{i, \rm syn}(t) = -I_{i, \rm syn}(t)/\tau_{\rm syn} 
\end{equation}
\noindent where $\tau_{\rm syn}$ is the postsynaptic current time constant (parameter values are shown in Table \ref{table:parameters}).

\begin{table}[!ht]
  \caption{\label{table:parameters} Neuron and synaptic parameters. Neuron and synaptic parameters used in our simulations according to \cite{potjans2014}.}
 \begin{center}
  \begin{tabular}{lccccccccc}
   \toprule
   \midrule
 membrane time constant & $\tau_m$ & 10 ms \tabularnewline
 refractory period & $\tau_{ref}$ & 2 ms \tabularnewline
 postsynaptic current time constant & $\tau_{syn}$ & 0.5 ms \tabularnewline
 membrane capacitance & $C_m$ & 250 pF\tabularnewline
 reset voltage & $V_{\rm reset}$ & -65 mV\tabularnewline
 threshold voltage & $V_{\rm th}$ & -50 mV\tabularnewline
 excitatory synaptic weight & $w$ & $\mathcal{N}$($\mu = 87.8, \sigma = 8.8$) pA \tabularnewline
 relative inhibitory synaptic weight & $g$ & 4 \tabularnewline
 excitatory synaptic transmission delay & $d_e$ & $\mathcal{N}$($\mu=1.5 , \sigma = 0.75$) ms \tabularnewline
 inhibitory synaptic transmission delay & $d_i$ & $\mathcal{N}$($\mu = 0.8 , \sigma = 0.4$) ms \tabularnewline
 initial membrane potential & $V_0$ & $\mathcal{N}$($\mu=-65.0, \sigma=6.5$) mV \tabularnewline
   \bottomrule
  \end{tabular}
 \end{center}
\end{table}

@@

\begin{table}[!ht]
  \caption{\label{table:population} Layer population sizes (extracted from \cite{potjans2014}). Neurons were distributed over four different layers (L23, L4, L5 and L6), and for each layer they were divided into excitatory (L23e, L4e, L5e and L6e) and inhibitory (L23i, L4i, L5i and L6i) subpopulations.}
 \begin{center}
  \begin{tabular}{cccccccc}
   \toprule
   \midrule
 L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i\tabularnewline
 20683 & 5834 & 21915 & 5479 & 4850 & 1065 & 14395 & 2948\tabularnewline
   \bottomrule
  \end{tabular}
 \end{center}
\end{table}

\begin{table}[!ht]
\caption{\label{table:conectivity} Connectivity matrix between the different populations of the model (extracted from \cite{potjans2014}). The connectivity matrix describes the probabilities of the target-specific connections between populations of neurons.
}
\begin{center}
\begin{tabular}{lccccccccc}
 \toprule
 \midrule
 &  &  &  &  & from &  &  &  & \tabularnewline
  \midrule
 &  & L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i\tabularnewline
 \midrule
 & L23e & 0.101 & 0.169 & 0.044 & 0.082 & 0.032 & 0 & 0.008 & 0\tabularnewline
 & L23i & 0.135 & 0.137 & 0.032 & 0.052 & 0.075 & 0 & 0.004 & 0\tabularnewline
 & L4e & 0.008 & 0.006 & 0.050 & 0.135 & 0.007 & 0.0003 & 0.045 & 0\tabularnewline
 to & L4i & 0.069 & 0.003 & 0.079 & 0.160 & 0.003 & 0 & 0.106 & 0\tabularnewline
 & L5e & 0.100 & 0.062 & 0.051 & 0.006 & 0.083 & 0.373 & 0.020 & 0\tabularnewline
 & L5i & 0.055 & 0.027 & 0.026 & 0.002 & 0.060 & 0.316 & 0.009 & 0\tabularnewline
 & L6e & 0.016 & 0.007 & 0.021 & 0.017 & 0.057 & 0.020 & 0.040 & 0.225\tabularnewline
 & L6i & 0.036 & 0.001 & 0.003 & 0.001 & 0.028 & 0.008 & 0.066 & 0.144\tabularnewline
\bottomrule
\end{tabular}
\end{center}
\end{table}

The ordinary differential equations were solved with the exact integration method for linear equations available in Brian 2 with a time step $\Delta t=0.1$ ms. All simulations were carried out in a cluster with 4 nodes each equipped with 2 Intel Xeon processors.

@@

\begin{enumerate}
\item \textbf{Layer specific}: Neurons from each layer receive specific background spike-trains drawn from an $8$ Hz Poisson distribution. The number of inputs per neuron is given in the first row of Table \ref{table:externalinput}.
\item \textbf{Layer independent}: Spike-trains are drawn from an $8$ Hz Poisson distribution as above, but now the number of inputs per neuron is the same for all the excitatory layers and the same for all the inhibitory layers as shown in the second row of Table \ref{table:externalinput}.
\item \textbf{DC input}: The Poissonian background is replaced by constant DC currents to all neurons. The number of inputs per neuron follows the layer specific configuration (first row in Table \ref{table:externalinput}). Observe that the number of inputs per neuron is multiplied by an effective factor which is given by $\nu \times \omega \times \tau_{syn}=0.3512$ pA. 
\end{enumerate}

@@

\begin{equation}
\label{Eq:Ksyn_alt}
C_{i} = \left( \frac{1-T}{1+T} \right)C_{e} ,
\label{Eq:transitivity}
\end{equation}

@@

To compare the distributions of firing rates and CVs obtained from simulations in Brian 2 and NEST we use the Kolmogorov-Smirnov statistical test. To apply this non-parametric test, cumulated histograms were constructed with bins chosen by the Doane method \cite{doane1976}.

\section{Results}\label{results}

@@

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.8]{fig2.pdf}
 \caption{Spontaneous cell-type specific activity (to be compared with Fig.6 from the original article). In this simulation, $g=4$ and background rate is layer-specific ($8$ Hz). (A) Raster plot of spiking activity for 0.4 s of all cell layers (from top to bottom; dark color: Excitatory cells, light color: Inhibitory cells). Number of displayed spike trains corresponds to 2.5\% of the total number of neurons (preserving relative number of cells per layer). (B–D) Statistics based on samples of 1000 neurons per layer recorded for 60 s. (B) Single-cell firing rate boxplot. Triangles indicate population mean firing rates, crosses show outliers. (C) Irregularity of single-cell firing rates. (D) Synchrony of spiking activity.}\label{fig:fig2}
\end{figure}

@@

@@

@@

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.02]{fig3.pdf}
 \caption{Cumulative distribution of CVs of ISIs for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange). NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig3}
\end{figure}

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.01]{fig4.pdf}
 \caption{Cumulative distribution of firing rate for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange). NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig4}
\end{figure}

When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, most of the layer-specific comparisons yielded p-values higher than 0.6 as before, indicating no significant differences. However, the p-value of the comparison between the CV of ISIs for the layer 5 excitatory neurons was $0.07$, and the p-values of the comparisons between the firing rates of excitatory cells from layers 2/3 and 6 were, respectively, $0.10$ and $0.08$. Therefore, in the cases of excitatory cells of layers 2/3, 5 and 6 there were statistically significant differences in the firing properties of neurons. The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have firing rate 12.2 $\pm$ 6.1 Hz while in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.

@@

@@

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.01]{fig5.pdf}
 \caption{Raster plot of spiking activity over a 0.4 s period and average firing rate for different types of external input (to be compared with Fig. 7 of the original article). Color codes and sample cell sizes as in Fig. \rot{\ref{fig:fig2}}. (A1--A2) DC input. (B1--B2) Layer independent input. (C1--C4) Histograms containing the population firing rates of 100 trials where the external inputs were drawn with a specific rule (see details in methods). }\label{fig:fig5}
\end{figure}

@@

@@

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.01]{fig6.pdf}
 \caption{Network activity dependence on background rate and relative inhibitory weight (to be compared with Fig. 8 of the original article). (A) Mean population firing rates of excitatory neurons in layers 2/3 (dashed line), 4 (triangles), 5 (stars) and 6 (squares) as a function of the background rate for fixed $g=4$. (B) AIness\% (see Methods) as a function of the background rate and the relative inhibitory synaptic strength $g$. Labeled dashed contour lines indicate areas where 25\%, 50\% and 75\% of all populations fire in an asynchronous and irregular mode at low rate. (C) Mean population firing rates of excitatory neurons as a function of the relative inhibitory synaptic strength $g$ for fixed background rate $8$ Hz (markers as in (A)).}\label{fig:fig6}
\end{figure}

@@

The successful replication of the results of the reference article confirms the correctness of the original implementation of the model.

\section{Acknowledgements}\label{acknowledgements}
The authors would like to thank Prof. Markus Diesmann for valuable comments on the manuscript. This research was supported by the S\~{a}o Paulo State Research Foundation (FAPESP) grant No.2013/07699-0 for the Research, Innovation and Dissemination Center for Neuromathematics, FAPESP grant 2016/03855-5 to NLK, FAPESP scholarship 2017/07688-9 to RS, FAPESP scholarship 2013/25667-8 to RP and FAPESP scholarship 2017/05874-0 to VC. ACR is partially supported by the CNPq fellowship grant 306251/2014-0. RP and ACR are also part of the IRTG 1740 / TRP 2015/50122-0, funded by DFG / FAPESP. CCC, and CR are supported by a CAPES PhD scholarship.

>>>>

\documentclass[10pt,a4paper,onecolumn]{article}
% * <aroquesilva@gmail.com> 2017-09-19T13:55:25.418Z:
%
% ^.
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\usepackage{cleveref} %added by Renan
\usepackage{multirow} %added by Rodrigo
\usepackage{amsmath,amssymb} %added by Roque
\usepackage[normalem]{ulem} % added by Nilton to use \sout
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Renan O. Shimoura,
      Nilton L. Kamiji,
      Rodrigo F.O. Pena,
  },
            pdftitle=
{
[Re] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}

@@

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017}

@@

Network neurons are described by the leaky integrate-and-fire (LIF) neuron model. The subthreshold membrane voltage of neuron $i$ obeys the equation
\begin{equation}
\label{Eq:LIF}
\dot{V}_i(t) = - \left( V_i(t) - V_{\rm reset} \right)/\tau_m + I_i(t)/C_m,
\end{equation}
where $\tau_m$ is the membrane time constant, $C_m$ is the membrane capacitance, $V_{\rm reset}$ is the reset potential, and $I_i(t)$ is the total input current. When $V_i(t) \geq V_{\rm th}$ the neuron emits a spike and the voltage is reset to $V_{\rm reset}$, remaining fixed at $V_{\rm reset}$ for a refractory period $\tau_{\rm ref}$. The total input current $I_i(t)$ is divided into external $I_{i, \rm ext}(t)$ and synaptic $I_{i, \rm syn}(t)$. Whenever an excitatory (or inhibitory) neuron $j$ presynaptic to neuron $i$ fires at time $t_{j}^{f}$, the synaptic current to neuron $i$ changes at time $t_{j}^{f}+d_e$ (or $d_i$) by an amount $w$ (or $-g w$), where $w$ is the excitatory synaptic weight, $g$ is the relative weight of the inhibitory synapse, $d_e$ is the excitatory synapse transmission delay, and $d_i$ is the inhibitory synapse transmission delay. In the absence of synaptic inputs the synaptic current changes as
\begin{equation}
\label{Eq:Isyn}
\dot{I}_{i, \rm syn}(t) = -I_{i, \rm syn}(t)/\tau_{\rm syn} 
\end{equation}
\noindent where $\tau_{\rm syn}$ is the postsynaptic current time constant (parameter values are shown in Table \ref{table:parameters}).

\begin{table}[!ht]
  \caption{\label{table:parameters} Neuron and synaptic parameters. Neuron and synaptic parameters used in our simulations according to \cite{potjans2014}.}
 \begin{center}
  \begin{tabular}{lccccccccc}
   \toprule
   \midrule
 membrane time constant & $\tau_m$ & 10 ms \tabularnewline
 refractory period & $\tau_{ref}$ & 2 ms \tabularnewline
 postsynaptic current time constant & $\tau_{syn}$ & 0.5 ms \tabularnewline
 membrane capacitance & $C_m$ & 250 pF\tabularnewline
 reset voltage & $V_{\rm reset}$ & -65 mV\tabularnewline
 threshold voltage & $V_{\rm th}$ & -50 mV\tabularnewline
 excitatory synaptic weight & $w$ & $\mathcal{N}$($\mu = 87.8, \sigma = 8.8$) pA \tabularnewline
 relative inhibitory synaptic weight & $g$ & 4 \tabularnewline
 excitatory synaptic transmission delay & $d_e$ & $\mathcal{N}$($\mu=1.5 , \sigma = 0.75$) ms \tabularnewline
 inhibitory synaptic transmission delay & $d_i$ & $\mathcal{N}$($\mu = 0.8 , \sigma = 0.4$) ms \tabularnewline
 initial membrane potential & $V_0$ & $\mathcal{N}$($\mu=-58.0, \sigma=10.0$) mV \tabularnewline
   \bottomrule
  \end{tabular}
 \end{center}
\end{table}

@@

\begin{table}[!ht]
  \caption{\label{table:population} Layer population sizes (extracted from \cite{potjans2014}). Neurons were distributed over four different layers (L23, L4, L5 and L6), and for each layer they were divided into excitatory (L23e, L4e, L5e and L6e) and inhibitory (L23i, L4i, L5i and L6i) subpopulations. The thalamic population is represented by Th.}
 \begin{center}
  \begin{tabular}{ccccccccc}
   \toprule
   \midrule
 L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i & Th\tabularnewline
 20683 & 5834 & 21915 & 5479 & 4850 & 1065 & 14395 & 2948 & 902\tabularnewline
   \bottomrule
  \end{tabular}
 \end{center}
\end{table}

\begin{table}[!ht]
\caption{\label{table:conectivity} Connectivity matrix between the different populations of the model (extracted from \cite{potjans2014}). The connectivity matrix describes the probabilities of the target-specific connections between populations of neurons.
}
\begin{center}
\begin{tabular}{lcccccccccc}
 \toprule
 \midrule
 &  &  &  &  & from &  &  &  & \tabularnewline
  \midrule
 &  & L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i & Th\tabularnewline
 \midrule
 & L23e & 0.101 & 0.169 & 0.044 & 0.082 & 0.032 & 0.0 & 0.008 & 0.0&0.0\tabularnewline
 & L23i & 0.135 & 0.137 & 0.032 & 0.052 & 0.075 & 0.0 & 0.004 & 0.0&0.0\tabularnewline
 & L4e & 0.008 & 0.006 & 0.050 & 0.135 & 0.007 & 0.0003 & 0.045 & 0.0&0.0983\tabularnewline
 to & L4i & 0.069 & 0.003 & 0.079 & 0.160 & 0.003 & 0.0 & 0.106 & 0.0&0.0619\tabularnewline
 & L5e & 0.100 & 0.062 & 0.051 & 0.006 & 0.083 & 0.373 & 0.020 & 0.0&0.0\tabularnewline
 & L5i & 0.055 & 0.027 & 0.026 & 0.002 & 0.060 & 0.316 & 0.009 & 0.0&0.0\tabularnewline
 & L6e & 0.016 & 0.007 & 0.021 & 0.017 & 0.057 & 0.020 & 0.040 & 0.225&0.0512\tabularnewline
 & L6i & 0.036 & 0.001 & 0.003 & 0.001 & 0.028 & 0.008 & 0.066 & 0.144&0.0196\tabularnewline
\bottomrule
\end{tabular}
\end{center}
\end{table}

The ordinary differential equations were solved with the exact integration method for linear equations available in Brian 2 with a time step $\Delta t=0.1$ ms. All simulations were carried out in a computer equipped with 2 Intel Xeon processors (E-2650v3, 2.3GHz) with 128GB of RAM.

@@

\begin{enumerate}
\item \textbf{Layer specific}: Neurons from each layer receive specific background spike-trains drawn from a Poisson distribution with firing rate $\nu = 8$Hz. The number of inputs per neuron is given in the first row of Table \ref{table:externalinput}.
\item \textbf{Layer independent}: Spike-trains are drawn from an $8$ Hz Poisson distribution as above, but now the number of inputs per neuron is the same for all the excitatory layers and the same for all the inhibitory layers as shown in the second row of Table \ref{table:externalinput}.
\item \textbf{DC input}: The Poissonian background is replaced by constant DC currents to all neurons. The number of inputs per neuron follows the layer specific configuration (first row in Table \ref{table:externalinput}). Observe that the number of inputs per neuron is multiplied by an effective factor which is given by $\nu \times w \times \tau_{syn}=0.3512$ pA. This factor is determined by calculating the time average of the synaptic current delivered by $n$ spikes of the Poisson background to a given neuron.
\item \textbf{Transient thalamic input}: A thalamic population (Th) of $902$ neurons, represented by Poissonian spike trains with firing rate of $120$ Hz, is connected with the network following the connection probabilities in Table \ref{table:conectivity}. In a simulated time period of $1$ s, the thalamic input is applied at time $t = 700$ ms and lasts for $10$ ms. This protocol is repeated hundred times to compute the average activity of $2.5$\% randomly sampled neurons from each layer of the network.
\end{enumerate}

@@

\begin{equation}
  \label{Eq:transitivity}
  C_{i} = \left( \frac{1-T}{1+T} \right)C_{e} ,
\end{equation}

@@

To compare the distributions of firing rates and CVs obtained from simulations in Brian 2 and NEST we use the Kolmogorov-Smirnov statistical test (KS-statistics). To apply this non-parametric test, cumulated histograms were constructed with bins chosen by the Doane method \cite{doane1976}. The null hypothesis that two sample sets were drawn from the same distribution is accepted, for an error $\alpha = 0.01$, if the KS-statistic $\leq 1.63$.
\section{Results}\label{results}

@@

\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.001]{fig2.pdf}
 \caption{Spontaneous cell-type specific activity (to be compared with Fig.6 from the original article). In this simulation, $g=4$ and background rate is layer-specific ($8$ Hz). (A) Raster plot of spiking activity for 0.4 s of all cell layers (from top to bottom; dark color: Excitatory cells, light color: Inhibitory cells). Number of displayed spike trains corresponds to 2.5\% of the total number of neurons (preserving relative number of cells per layer). (B–D) Statistics based on samples of 1000 neurons per layer recorded for 60 s. (B) Single-cell firing rate boxplot. Triangles indicate population mean firing rates, crosses show outliers. (C) Irregularity of single-cell firing rates. (D) Synchrony of spiking activity.}\label{fig:fig2}
\end{figure}

@@

\begin{figure}[!htp]
 \centering
 \includegraphics[scale=0.01]{fig3.pdf}
 \caption{Cumulative distribution of CVs of ISIs for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange). NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig3}
\end{figure}

@@

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.005]{fig4.pdf}
 \caption{Cumulative distribution of firing rate for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange). NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig4}
\end{figure}

@@

When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, all layer-specific comparisons yielded KS-statistics lower than 1.14, indicating no significant differences for an error $\alpha = 0.01$. The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have mean firing rate 12.2 $\pm$ 6.1 Hz, while using equation~(\ref{Eq:Ksyn}) their mean firing rate is 7.8 $\pm$ 4.8 Hz and in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.

@@

\begin{figure}[!htb]
 \centering
 \includegraphics[scale=0.005]{fig5.pdf}
 \caption{Raster plot of spiking activity over a 0.4 s period and average firing rate for different types of external input (to be compared with Fig. 7 of the original article). Color codes and sample cell sizes as in Fig. \rot{\ref{fig:fig2}}. (A1--A2) DC input. (B1--B2) Layer independent input. (C1--C4) Histograms containing the population firing rates of 100 trials where the external inputs were drawn with a specific rule (see details in methods). }\label{fig:fig5}
\end{figure}

@@

@@

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.1]{fig6.pdf}
 \caption{Network activity dependence on background rate and relative inhibitory weight (to be compared with Fig. 8 of the original article). (A) Mean population firing rates of excitatory neurons in layers 2/3 (dashed line), 4 (triangles), 5 (stars) and 6 (squares) as a function of the background rate for fixed $g=4$. (B) AIness\% (see Methods) as a function of the background rate and the relative inhibitory synaptic strength $g$. Labeled dashed contour lines indicate areas where 25\%, 50\% and 75\% of all populations fire in an asynchronous and irregular mode at low rate. (C) Mean population firing rates of excitatory neurons as a function of the relative inhibitory synaptic strength $g$ for fixed background rate $8$ Hz (markers as in (A)).}\label{fig:fig6}
\end{figure}

@@

\subsection{Propagation of Transient Thalamic Inputs}

In this section we make use of the thalamic inputs as described in the last column of Table \ref{table:population}. For further information on how thalamic inputs are applied we refer to the Methods section.

As described in the results section of the original article, when thalamic input is delivered to the network the output is rather stereotypical. This is confirmed by our results presented in Fig.~\ref{fig:fig7}, which shows the response of the network replicated in Brian2 to transient thalamic input. By comparing histograms in Fig.~\ref{fig:fig7}\textbf{B} with the ones in Fig. 10 \textbf{B} of the original article one can see the similarities.

\begin{figure}[!ht]
 \centering
% * <aroquesilva@gmail.com> 2018-02-06T01:02:54.595Z:
%
% ^.
 \includegraphics[scale=0.01]{fig7.pdf}
 \caption{Response to transient thalamic input (to be compared with panels \textbf{A} and \textbf{B} of Fig.10 from the original article). Note that the time axis is centralized at the beginning of the thalamic input, which lasts for 10 ms. (A) Sample raster plot of the network while receiving transient thalamic input as described in the Methods section. (B) Cell-type specific population spike-counts averaged over 100 different applications of thalamic stimuli.}\label{fig:fig7}
\end{figure}

\subsection{Important information needed during replication}

Most of the information necessary for the replication done here was available in the original article. However, a few details were found only in the NEST implementation of the code available in the Open Source Brain (OSB) platform \cite{potjans2014pyNEST}. In this section we list these details.

\begin{itemize}

\item When replacing the Poissonian background by a DC current the number of inputs per neuron was multiplied by the factor $0.3512$ pA as explained in the Methods section of this article. This factor is not explicitly explained in the original article.

\item There is no mention to the initial voltage values in the original article. This information was found in the OSB NEST code where one can observe that the initial voltage values are randomly drawn from the normal distribution $\mathcal{N}$($\mu=-58 , \sigma = 10$) mV.

\item In the original article, the equation for the integrate-and-fire model does not explicitly state the  resting potential. This could suggest to a na\"ive reader that the resting potential is $0$ mV, whereas the correct resting potential is the reset voltage $V_{\rm reset}$ (cf. equation~(\ref{Eq:LIF})).

\item The thalamic input in the original article is described as having $15$ Hz, however in the OSB NEST code it is actually $120$ Hz, which is the correct value to be used.

\end{itemize}

@@

The successful replication of the results of the reference article confirms the robustness of the original implementation of the model.

\section{Acknowledgements}\label{acknowledgements}
The authors would like to thank Prof. Markus Diesmann, Dr. Maximilian Schmidt and Dr. Sacha van Albada for valuable comments on the manuscript. This research was supported by the S\~{a}o Paulo State Research Foundation (FAPESP) grant No.2013/07699-0 for the Research, Innovation and Dissemination Center for Neuromathematics, FAPESP grant 2016/03855-5 to NLK, FAPESP scholarship 2017/07688-9 to RS, FAPESP scholarship 2013/25667-8 to RP and FAPESP scholarship 2017/05874-0 to VC. ACR is partially supported by the CNPq fellowship grant 306251/2014-0. RP and ACR are also part of the IRTG 1740 / TRP 2015/50122-0, funded by DFG / FAPESP. CCC, and CR are supported by a CAPES PhD scholarship.
